---
title: "Viz part 2"
author: "Sarah Haberlack"
date: "2023-10-03"
output: github_document
---

## Settings

```{r}
library(tidyverse)
library(patchwork)

theme_set(theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


Get the data for plotting today.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

This results in a dataframe with `r nrow(weather_df)` observations on 6 variables.

## Same plot from last time
Adding labels ("labs")

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (째C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  )
```

Now, let's play with scale. Default is good, but there are other options that you might sometimes want to use instead. Scales modify the default mapping between the variable and the aesthetic it's mapped onto.
Note: We can also "sqrt" transform the data using `trans = "sqrt"` within the `scale_y_continuous` command. It is sometimes easier to look at some data if it's transformed. You could also do this in the mutate function if you want.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (째C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right",
    limits = c(0,30)
  )
```

what about colors ...
for colors, continuous color options and discrete color options.
for categorical variables (not ordinal), `hue` works. but it's gross!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (째C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```


## Themes

`theme_bw()` inverts the background grid colors. theme_bw is a RESET theme command. a global preference. order matters here, theme_bw has to be first.
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (째C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_bw() +
theme(legend.position = "bottom")
```

## data argument...

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax,)) +
  geom_point(aes(color = name)) +
  geom_smooth()

nyc_weather_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

hawaii_weather_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = hawaii_weather_df)
```

## `patchwork`

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name)
```


```{r}
ggp_temp_scatter =
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

ggp_prcp_density =
  weather_df |> 
  filter(prcp > 25) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
    theme(legend.position = "bottom")

ggp_temp_scatter + ggp_prcp_density

ggp_tmax_date =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

(ggp_temp_scatter + ggp_prcp_density)/ggp_tmax_date
```

## data manipulation

```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))
  ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Factor variables are ordinal.
When ggplot thinks about orders, every time you need to put sth that was a character variable into a plot, in the background R categorizes them as factor variables by alphabetical order. If you want these things in a different order, you need to do data manipulation.

What if you want to order a variable in the order of a certain other variable? Let's order `name  ` by `tmax`.
```{r}
weather_df |> 
  mutate(
    name = fct_reorder(name, tmax)
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```
So now this is a somewhat more natural ordering, we can see how the data stacks up.

## complicated FAS plot
```{r}
litters_df =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), se = 3)
  
pups_df =
  read_csv("data/FAS_pups.csv") |> 
  janitor::clean_names()
  
fas_df =
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
  pd_ears:pd_walk,
  names_to = "outcome",
  values_to = "pn_day"
) |> 
drop_na() |> 
mutate(outcome = fct_reorder(outcome, pn_day)) |> 
ggplot(aes(x = dose, y = pn_day)) +
geom_violin() +
facet_grid(day_of_tx ~ outcome)
```

 
