---
title: "Viz part 2"
author: "Sarah Haberlack"
date: "2023-10-03"
output: github_document
---

```{r}
library(tidyverse)
```

```{r}
knitr::opts_chunk$set(
  fig.width = .6,
  fig.asp = .6,
  out.width = "90%"
)
```


Get the data for plotting today.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Same plot from last time
Adding labels ("labs")

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (°C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  )
```

Now, let's play with scale. Default is good, but there are other options that you might sometimes want to use instead. Scales modify the default mapping between the variable and the aesthetic it's mapped onto.
Note: We can also "sqrt" transform the data using `trans = "sqrt"` within the `scale_y_continuous` command. It is sometimes easier to look at some data if it's transformed. You could also do this in the mutate function if you want.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (°C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right",
    limits = c(0,30)
  )
```

what about colors ...
for colors, continuous color options and discrete color options.
for categorical variables (not ordinal), `hue` works. but it's gross!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (°C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations, data from rnoaa"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```


## Themes

