---
title: "Viz and EDA 1"
author: "Sarah Haberlack"
date: "2023-09-28"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

knitr::opts_chunk$set(
  fig.width = .6,
  fig.asp = .6,
  out.width = "90%"
)

Get the data for plotting today.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Let's make a plot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Pipes and stuff
sh: you're saying "start with this df, then do something else"
sh: save as a new df to condense your code, or to save plots.

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather =
  weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Fancy plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth()
```

Changing where the color is. In the below graph, only the points will be highlighted by color and the smooth line is not.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Taking away the points and just keeping the smooth line (practically, not a great idea but it is possible to do.)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_smooth(se = FALSE)
```

Only in geom point, make 70% opaque and 30% solid.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha= 0.3) +
  geom_smooth()
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

^ dot says "put everything on this row" and facet by name.
automatically outputs based on alphabetical order of facets.

Let's try a different plot. Temps are boring. Over that two-year period, what's the average temperature?

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

Try assigning a specific color (JG almost never sets color by hand in this way. )
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "blue", size = .5)  
```

hex plot
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

line plot
```{r}
weather_df |>
  filter(name == "Molokai_HI") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```


## Univariate plotting

Histogram

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

dodge bames the bars "avoid" each other instead of stacking on top of each other. but this looks confusing. to compare all this data, we'd rather just not use a histogram anymore to convey this data.

let's use a density plot.

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3, adjust = .5)
```

In a density plot, 
adjust" makes everything a little LESS smooth < 1 and more smooth > 1.

We can compare univariate distributions also by using BOXPLOTS.

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_boxplot()
```

Violin plots?

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_violin()
```

ridge plot

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
```

^Quick way of looking at, e.g., 50 states at a time

## saving and embedding plots

```{r}
ggp_weather =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather

ggsave("results/ggp_weather.pdf", ggp_weather)
```

When we knit, the plots become embedded into this document. We ask, is this the right size? etc? R is making automatic rescaling choices.

To change the scale of the plot being made,:

```{r, fig.width = 12, fig.asp = .6}
ggp_weather
```

fig.asp is "aspect ratio" in relation to the width